---
title: "exams2moodle"
author: 
- name: "Sigbert Klinke" 
  email: sigbert@hu-berlin.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
vignette: > 
  %\VignetteIndexEntry{exams2moodle} 
  %\VignetteEngine{knitr::rmarkdown} 
  \usepackage[utf8]{inputenc} 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Note**: Part the text is taken from the [Moodle](https://moodle.org/) website, the [`exams`](https://CRAN.R-project.org/package=exams) package and the [R/exams](http://www.r-exams.org/) web site. From my point of view there is no need to reinvent the wheel (or text).

Moodle is a learning platform designed to provide educators, administrators and learners with a single robust, secure and integrated system to create personalised learning environments. Part of Moodle is an activity Quiz which allows a the teacher to design and set quizzes consisting of a large variety of Question types. The questions are stored in a  Question bank and can be reused in different quizzes. 

The `exams` package in R allows the automatic generation of exams based on exercises in Markdown or LaTeX format, possibly including R code for dynamic generation of exercise elements. Exercise types include single-choice and multiple-choice questions, arithmetic problems, string questions, and combinations thereof (cloze). Output formats include standalone files (PDF, HTML, Docx, ODT, ...), **Moodle XML**, QTI 1.2, QTI 2.1, Blackboard, Canvas, OpenOLAT, ARSnova, and TCExam. 

Due to Corona I, as statistics teacher, was forced to create online exams. It was a natural way for me to create exercises with R Sweave (or R Markdown) for the Question bank in Moodle. But an online exam is in various ways different to a written exam. Since I'am more familiar with R then with Moodle I decided to create with the `exams` package dynamic questions, create `n` replicates using different numbers and upload the questions into Moodle.

To create exercises I use several steps:

* `exams2pdf(exercisename, nsamp=10)` to create **one** PDF file to check for possible problems in the questions or solutions,
* `stress_test(exercisename)` to see if any problems appear with the questions and to get an idea of the distribution of the correct answers, and 
* `exams2moodle(exercisename, n=10)` to create a Moodle XML file, here with ten replications. 

The difference of using either `n` or `nsamp` are of practical nature:

* In `exams2pdf` using `n=10` creates ten PDF files, one for each question,  whereas `nsamp=10` creates one PDF file with all questions.
* In `exams2moodle` using `n=10` creates one category with ten questions, whereas `nsamp=10` creates ten categories containing one question. Whenever you have to delete categories and questions in the question bank in Moodle then you will note the difference :)

## Question types

The `exams` package supports the following answer types for a question:

| `exams`              | Moodle          |
|----------------------|-----------------| 
| `mchoice`            | Multiple choice (multiple answers possible) | 
| `schoice`            | Multiple choice (only one answer possible)  | 
| `num`                | Numerical                                   | 
| `string`             | Short Answer                                |
| `essay`              | Essay                                       |
| `cloze`              | Embedded Answers (Cloze Test / Gap Fill)    |

Other Moodle answer types are not (yet?) supported: 

* Calculated ...
* Drag and drop ...
* Description
* (Random short-answer) Matching
* Select missing words 

Other could be emulated:

* True/False (`cloze` and `schoice`)

### Essay

```{r, eval=FALSE}
%% \extype{string}
%% \exsolution{nil}
%% \exextra[essay,logical]{TRUE}
%% \exextra[essay_format,character]{editor}
%% \exextra[essay_required,logical]{FALSE}
%% \exextra[essay_fieldlines,numeric]{5}
%% \exextra[essay_attachments,numeric]{1}
%% \exextra[essay_attachmentsrequired,logical]{FALSE}
%% \exmaxchars{1000, 10, 50}
```

## Frequently asked questions

### How can I force a block of LaTeX to become an image in Moodle?

Source: https://stackoverflow.com/questions/62222053/different-copies-of-question-with-table-for-moodle-with-r-exams

Use the command `tex2image` from the `exams` package:

```{r, eval=FALSE}
<<echo=FALSE, results=hide>>=
tab <- '\\begin{tabular}{lrr}
\\hline
Name & Min & Max \\\\
\\hline
Foo  & 0   & 1   \\\\
Bar  & 0   & 100 \\\\
\\hline
\\end{tabular}'
tex2image(tab, name = "tab", dir = ".", pt = 8, resize = 250)
@
\includegraphics{tab.png}
```

## Exercise types in `exams` and `Moodle`

```{r, echo=FALSE}
library("exams")
exerdir   <- paste0(system.file(package="exams"), "/exercises")
exerfiles <- list.files(path=exerdir, pattern="*.Rnw$")
extype    <- rep('', length(exerfiles))
for (i in seq(exerfiles)) {
  fcont     <- readLines(paste0(exerdir, '/', exerfiles[i]))
  extype[i] <- fcont[grepl("\\extype", fcont, fixed=TRUE)]
}
extype <- gsub('[ \\{\\}%]', '', extype)
extype <- gsub('extype', '', extype, fixed=TRUE)
#
tdir <- tempdir(TRUE)
res  <- exams2moodle(exerfiles, n=1, edir=exerdir, tdir=tdir, 
                     name='rexams', schoice = list(eval = exams_eval(rule = "none")))
library("XML")
xmlfile <- xmlParse("rexams.xml")
xmltop  <- xmlRoot(xmlfile)
xmlquestion <- xmlSApply(xmltop, xmlAttrs)
xmlquestion <- xmlquestion[xmlquestion!='category']
#
df <- data.frame(Example_exercise          = sapply(res[[1]], function(e) { e$metainfo$file}),
                 Exercise_type = extype,
                 Cloze_type    = sapply(res[[1]], function(e) { if (is.null(e$metainfo$clozetype)) return('');
                                                               return(paste0(e$metainfo$clozetype, collapse="|"))}),
                 Moodle_type   = xmlquestion,
                 stringsAsFactors = FALSE)
row.names(df) <- NULL
df <- df[order(df$Exercise_type, df$Moodle_type),]
library("knitr")
kable(df, row.names=FALSE)
```
